Finding Code Clone Refactoring Techniques by Mapping Clone Context

### 2021-11-10 cefp/paper.tex
Updates from Overleaf

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\usepackage[style=ieee,citestyle=numeric-comp,dashed=false]{biblatex}
\usepackage{booktabs}
\usepackage{multirow}
\addbibresource{refs.bib}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Finding Code Clone Refactoring Techniques by Mapping Clone Context}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Simon Baars\inst{1,2}\orcidID{0000-0001-7905-1027} \and
Ana Oprescu\inst{1,3}\orcidID{0000-0001-6376-0750}}
%
%\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Amsterdam \and
\email{simon.j.baars@gmail.com} \and
\email{A.M.Oprescu@uva.nl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Reducing clones in source code is one of the techniques to improve the maintainability of a software system. Which refactoring technique to use depends on where a clone is found and what the relation between clone instances in a clone class is. We define three influencing factors on how a clone should be refactored: Relation, Location and Contents. The Relation describes the inheritance relation among the clone instances in a clone class. The Location describes where a clone instance is found in the source code. The Contents describe what a clone instance spans.

Based on experiments on corpus of open-source Java projects we find that most clones (77\%) are in the body of methods or constructors and thus the ``Extract Method'' refactoring technique applies. What further techniques are required for the refactoring depends on the relation among the clone instances of a clone. We define four relations that require different further refactoring techniques: Common Class, Common Hierarchy, Common Interface and Unrelated. % (ordered by how favorable their refactoring is). When clones are in the same class, no further techniques than ``Extract Method'' have to be applied. If they are not, further techniques have to be applied that create a trade-off on system maintainability. 
The closer classes are related, the more favorable refactoring clones by such relations becomes. 37\% of clones are in the same class, 24\% share an inheritance hierarchy, 24\% are unrelated and 15\% have a common interface.% For each of these, we define subcategories to facilitate further understanding in what relations clones are found in the codebase.

% What further techniques apply when refactoring a clone using the ``Extract Method'' technique depends on the relation among its clone instances. If clone instances share a common class, no further refactoring techniques are required. If they are in the same inheritance hierarchy, ``Pull-Up Method'' can be used till the extracted method is in a location accessible by all instances.

\keywords{Code Clones \and MSR \and Clone Relation \and Inheritance \and Object-Oriented Programming.}
\end{abstract}
%
%
%
\section{Introduction}
Duplicate code fragments are often considered as bad design~\cite{fowler2018refactoring}. They increase maintenance efforts or cause bugs in evolving software~\cite{heitlager2007practical}. Changing one occurrence of a duplicated fragment may require changes in other occurrences~\cite{ostberg2014automatically}. Furthermore, duplicated code was shown to account for up to 25\% of total system volume~\cite{bruntink2005use}, entailing more code to be maintained.

@@

%We perform statistical measurements on a large corpus of open-source software systems to determine which relation, location and contents category has the most clones. We find that 77\% of clones is found in the bodies of constructors and methods, which indicate clones that can be refactored by applying the ``Extract Method'' refactoring technique. Clones were found in very diverse relations, which indicates that in
We perform statistical measurements on a large corpus of open-source software systems to determine which relation, location and contents category has the most clones. We find that a significant portion of clones is found in the bodies of constructors and methods, which indicate clones that can be refactored by applying the ``Extract Method'' refactoring technique. What further techniques apply when refactoring such a clone depends on the relation among its clone instances.

@@

This is shown schematically in Figure~\ref{fig:clonecontext}.
%We define categories for each of these aspects and determine the categories of clones.

@@

%Fig.~\ref{fig:clonecontext} shows an abstract representation of clone classes and clone instances. The relation of clones through inheritance is determined for each clone class. The location and contents of clones are determined for each clone instance.

@@

%When CloneRefactor applies automated refactorings, it uses this inheritance relation to see where it must place the refactored code. We explain this for each relation category over the following sections.

%To determine the relation for a given clone class, we compare each node in a clone instance with the respective nodes in each other clone instance within a clone class. %So for a clone class with x clone instances and each clone instance has y nodes, we perform:

%\begin{equation}\label{eq:sameclass}
%\frac{x * (x + 1)}{2} * y
%\end{equation}

%comparisons to determine the relation. Each of these checks will result in one of the relations from the list above. Of these relations, the item lowest in the list is going to be the relation of the clone class. So for instance, if a clone class has 15 nodes that denote a \textit{Superclass} relation but 3 nodes are \textit{Unrelated}, the clone class becomes \textit{Unrelated}.

@@

 %We define the method as the first method declaration that is encountered when following the parent nodes in the AST for a given cloned node. %Please note that a clone instance may not always be in a method, for which this predicate will fail.

%Clone classes are flagged as \textit{Same method} if all their clone instances are found in the same method. Cloned classes are flagged as \textit{Same class} if all their clone instances are found in the same class. %We define the class as the first class declaration that is encountered when following the parent nodes in the AST for a given cloned node.

@@

%Clone classes are flagged as \textit{Superclass} if the classes in which the clone instances are found are parent and child class of each other. Clone classes are flagged as \textit{Siblings} if the classes in which their clone instances are found all share the same parent and this parent class is not external. Clone classes are flagged as \textit{First Cousin} if the classes in which their clone instances are found all share the same grandparent and this grandparent class is not external.

%Clone classes nodes are flagged as \textit{Ancestor} if the classes in which their clone instances are found are recursively the parent of an antecedent. Cloned nodes are flagged as \textit{Same Hierarchy} if the classes in which these nodes are found are all in the same inheritance hierarchy and not linked by any external classes.

@@

%To check whether two cloned nodes have a \textit{Same Direct Interface} relation, we check that the classes to which these nodes belong both implement the same interface type. To check whether two cloned nodes have a \textit{Same Indirect Interface} relation, we recursively walk all types the class or interface of the cloned nodes implement and extend. The common interface that is closest to the cloned node in terms of depth of recursion will be flagged as the refactoring candidate.

@@

%Cloned nodes are flagged \textit{No Direct Superclass} if they are in classes that do currently not have a parent. This marks the opportunity for creating a superclass abstraction and placing the extracted method there. Cloned nodes are flagged \textit{No Indirect Superclass} if any of their ancestors do not have a parent. In such a case, it would be possible to create such an abstraction for the ancestor that does not have a parent.

%Cloned nodes are flagged \textit{External Superclass} if they are in a class which has an external parent. Cloned nodes are flagged \textit{External Ancestor} if one of their ancestors has an external parent. Both of these relations obstruct the possibility of creating a superclass abstraction. In such a case, an interface abstraction could be created to make their relation explicit.

@@

%To detect whether a given clone instance is at Method Level, we check for each node in the clone instance whether it has an ancestor node that is a method declaration. A clone instance is considered ``Method Level'' if:

%\begin{equation}\label{eq:samemethod}
%\forall (n \in I)\text{ } Tn = \text{MethodDeclaration} \lor \exists (p \in ancestors(n)) \text{ } Tp = \text{MethodDeclaration}
%\end{equation}

%Where \textit{T} is the type of a node. \textit{ancestors(n)} denotes all of ancestor nodes of a node. The ancestor nodes are all recursive parents of a given node in the AST. Apart from this, we check that the MethodDeclaration ancestor of each clone instance is the same, to be sure the clone only spans a \textit{single} method. Figure \ref{fig:methodlevelclone} shows an example of a method level clone and its recursive parents.

%\begin{figure}[H]
%\begin{parcolumns}{2}
%\colchunk[1]{
%\begin{javacode}
%  public void doA() { // |\textcolor{pgreen}{$p_2$}|
%    public class Class1 { // |\textcolor{pgreen}{$p_3$}|
%    if(isA()) // |\textcolor{pgreen}{$p_1$}|
%|\highlightYellow|      doB(); // |\textcolor{pgreen}{$n_1$}|
%  }
%}
%\end{javacode}}
%\colchunk[2]{
%\begin{javacode}
%public class Class2 { // |\textcolor{pgreen}{$p_3$}|
%  public void doB() { // |\textcolor{pgreen}{$p_2$}|
%    if(isB()) // |\textcolor{pgreen}{$p_1$}|
%|\highlightYellow|      doB(); // |\textcolor{pgreen}{$n_2$}|
%  }
%}
%\end{javacode}}
%\end{parcolumns}
%\caption{Example of a method level clone and its parents.}
%\label{fig:methodlevelclone}
%\end{figure}

@@

%We detect these clones in the same way we detect method level clones: we walk up the AST until we find the first node that indicates one of these categories.

@@

\subsubsection{Full Method/Constructor/Class/Interface/Enumeration}
These categories denote that a full declaration, including its body, is cloned with another declaration. These categories often denote redundancy and are often easy to refactor: one of both declarations is redundant and should be removed. All usages of the removed declaration should be redirected to the clone instance that was not removed. Sometimes, the declaration should be moved to a location that is accessible by all usages. %We check for clones in this category by checking that the first node in a clone instance is a declaration and the last node in the clone instance is the last node in the body of that declaration.

\subsubsection{Partial Method/Constructor}
These categories describe clone instances which are found in the body of a method or constructor. These clones can often be refactored by extracting a new method out of the cloned code. %We detect such clones by checking that each node in the clone instance has an ancestor node that is a MethodDeclaration, but none of the nodes is a method declaration. We describe the procedure by with we recursively seek the parent nodes of a node in Section~\ref{sec:methodlevelcr}.

\subsubsection{Several Methods}
Several methods cloned in a single class is a strong indication of implicit dependencies between two classes. This increases the chance that these classes are missing some form of abstraction, or their abstraction is used inadequately. %We detect such clones by checking that each cloned node has a parent node that is a MethodDeclaration. This is the same process as described in equation \ref{eq:samemethod}. Additionally, we check that the MethodDeclaration ancestor of the first node in the clone instance differs from the last, to be sure that the clone instance spans over at least two methods.

\subsubsection{Only Fields}
This category denotes that the clone spans over only global variables/fields that are declared outside of a method. This indicates data redundancy: pieces of data have an implicit dependency. In such cases, these fields may have to be encapsulated in a new object. Or, the fields should be somewhere in the inheritance structure where all objects containing the clone can access them. %We check for this category by validating that all nodes in the clone instance are a VariableDeclarator. Because VariableDeclarators in methods would already fall into the ``Partial Method'' category, this includes only globally defined fields.

@@

\subsection{Relation} \label{sec:relationresults}
Table~\ref{tab:relation} displays the number of clone classes found for the entire corpus for different relations (see Section~\ref{sec:setuprelation}).
% Sander: is dat plaatje met overzicht van relaties van clone classen weg of ergens anders? hij is hier ook wel handig

>>>>

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\usepackage[style=ieee,citestyle=numeric-comp,dashed=false]{biblatex}
\usepackage{booktabs}
\usepackage{multirow}
\addbibresource{refs.bib}

\begin{document}
\title{Finding Code Clone Refactoring Techniques by Mapping Clone Context}

\author{Simon Baars\inst{1,2}\orcidID{0000-0001-7905-1027} \and
Ana Oprescu\inst{1,3}\orcidID{0000-0001-6376-0750}}

\institute{University of Amsterdam \and
\email{simon.j.baars@gmail.com} \and
\email{A.M.Oprescu@uva.nl}}
\maketitle

\begin{abstract}
Reducing clones in source code is one of the techniques to improve the maintainability of a software system. Which refactoring technique to use depends on where a clone is found and what the relation between clone instances in a clone class is. We define three influencing factors on how a clone should be refactored: Relation, Location and Contents. The Relation describes the inheritance relation among the clone instances in a clone class. The Location describes where a clone instance is found in the source code. The Contents describe what a clone instance spans.

Based on experiments on corpus of open-source Java projects we find that most clones (77\%) are in the body of methods or constructors and thus the ``Extract Method'' refactoring technique applies. What further techniques are required for the refactoring depends on the relation among the clone instances of a clone. We define four relations that require different further refactoring techniques: Common Class, Common Hierarchy, Common Interface and Unrelated. The closer classes are related, the more favorable refactoring clones by such relations becomes. 37\% of clones are in the same class, 24\% share an inheritance hierarchy, 24\% are unrelated and 15\% have a common interface.

\keywords{Code Clones \and MSR \and Clone Relation \and Inheritance \and Object-Oriented Programming.}
\end{abstract}

\section{Introduction}
Duplicate code fragments are often considered as bad design~\cite{fowler2018refactoring}. They increase maintenance efforts or cause bugs in evolving software~\cite{heitlager2007practical}. Changing one occurrence of a duplicated fragment may require changes in other occurrences~\cite{ostberg2014automatically}. Furthermore, duplicated code was shown to account for up to 25\% of total system volume~\cite{bruntink2005use}, entailing more code to be maintained.

@@

We perform statistical measurements on a large corpus of open-source software systems to determine which relation, location and contents category has the most clones. We find that a significant portion of clones is found in the bodies of constructors and methods, which indicate clones that can be refactored by applying the ``Extract Method'' refactoring technique. What further techniques apply when refactoring such a clone depends on the relation among its clone instances.

@@

This is shown schematically in Figure~\ref{fig:clonecontext}.

@@

@@

@@

@@

@@

@@

@@

@@

@@

\subsubsection{Full Method/Constructor/Class/Interface/Enumeration}
These categories denote that a full declaration, including its body, is cloned with another declaration. These categories often denote redundancy and are often easy to refactor: one of both declarations is redundant and should be removed. All usages of the removed declaration should be redirected to the clone instance that was not removed. Sometimes, the declaration should be moved to a location that is accessible by all usages.

\subsubsection{Partial Method/Constructor}
These categories describe clone instances which are found in the body of a method or constructor. These clones can often be refactored by extracting a new method out of the cloned code.

\subsubsection{Several Methods}
Several methods cloned in a single class is a strong indication of implicit dependencies between two classes. This increases the chance that these classes are missing some form of abstraction, or their abstraction is used inadequately.

\subsubsection{Only Fields}
This category denotes that the clone spans over only global variables/fields that are declared outside of a method. This indicates data redundancy: pieces of data have an implicit dependency. In such cases, these fields may have to be encapsulated in a new object. Or, the fields should be somewhere in the inheritance structure where all objects containing the clone can access them.

@@

\subsection{Relation} \label{sec:relationresults}
Table~\ref{tab:relation} displays the number of clone classes found for the entire corpus for different relations (see Section~\ref{sec:setuprelation}).
